---
title: "Class 5 script"
output: html_document
---
```{r}

```

```{r setup, include=FALSE}
# Load libraries
library(tidyverse)
library(rpart)
library(rpart.plot)
```

## Agenda

1. Review portions of the last project.

2. Prepare for module 5 project.  I hope you have noticed that the code in the webinar can be adapted pretty directly for the project!

Avoid frustration! If you hit an obstacle in the project (which is making you feel like hitting something), please get in touch with us!

## Megatelco Data Dictionary 

 DEMOGRAPHIC VARIABLES:
 - College - has the customer attended some college (one, zero)
 - Income - annual income of customer
 - House - estimated price of the customer's home (if applicable)

 USAGE VARIABLES:
 - Data Overage Mb - Average number of megabytes that the customer used in excess of the plan limit (over last 12 months)
 - Data Leftover Mb - Average number of megabytes that the customer use was below the plan limit (over last 12 months)
 - Data Mb Used - Average number of megabytes used per month (over last 12 months)
 - Text Message Count - Average number of texts per month (over last 12 months)
 - Over 15 Minute Calls Per Month - Average number of calls over 15 minutes in duration per month (over last 12 months)
 - Average Call Duration- Average call duration (over last 12 months)

PHONE VARIABLES:
 - Operating System - Current operating system of phone
 - Handset Price - Retail price of the phone used by the customer

ATTITUDINAL VARIABLES:
 - Reported Satisfaction - Survey response to "How satisfied are you with your current phone plan?" (high, med, low)
 - Reported Usage Level - Survey response to "How much do your use your phone?" (high, med, low)
 - Considering Change of Plan - Survey response to "Are you currently planning to change companies when your contract expires?" (high, med, low)

OTHER VARIABLES
 - Leave - Did this customer churn with the last contract expiration? (LEAVE, STAY)
 - ID - Customer identifier

## Tips for compiling/knitting your RMarkdown document

The project for this module, like the last module, includes a file upload. You must knit the RMarkdown document in which you did your work.  

Key concept:  RMarkdown will only use the information in your code chunks as it compiles! This can create problems if you don't have all the info the document needs in the chunks! 

Solution: 

    1. Clear your environment.
    2. Run each chunk in sequence looking for errors.
    3. Fix errors.
    4. Repeat 1 and 2. Now if there are no errors you should be good to go.
    5. Compile.
    6. Submit the automatically generated word or HTML file (find it in your working directory) for the file upload assignment.

## Example Code for Project

What do you need to know for this module's project?

1. Fit a tree model using all the available predictors. 
2. Create a confusion matrix using predict() with the `type = "class"` argument and identify numbers of TP, FP, TN and FN.
3. Estimate profit (benefits - costs) using a defined cost-benefit matrix and the above confusion matrix.
4. Estimate profit for a strategy that consists in calling everyone.
5. Estimate profit for a class decision threshold of .3 using predict() with the `type = "prob"` argument.
6.  Use the model to predict on a new dataset (without the target), then use these predictions to identify those who should be called--a contact list.
7.  Make a recommendation to the Director of Sales based on your analytic work.

Note:  For the module 5 project you will predict answering for prospective customers (who have not yet been called). Consequently, the assignment requires an additional dataset, adviseinvest_new_customer.csv, that you will need to download and move into your project folder.  Here is code to load that data into memory after you have it saved in your project folder (your working directory). This code is included in the template script.

We will again use the MegaTelCo data for this demonstration.

```{r}
# Load and clean data--recycle code from last project demo

m <- read_csv(file = "https://raw.githubusercontent.com/Stan-Pugsley/mktg_6487_base/main/Assignments/DataSets/megatelco.csv")

m_clean <- m |> 
  mutate(reported_satisfaction = factor(reported_satisfaction,
                                        levels = c("low","avg", "high"),
                                        ordered = T),
         reported_usage_level = factor(reported_usage_level,
                                       levels = c("low", "avg", "high"),
                                       ordered = T),
         considering_change_of_plan = factor(considering_change_of_plan,
                                             levels = c("no", "maybe", "yes")),
         leave = factor(leave), # let the levels be assigned alphabetically
         college = ifelse(college=="one", "yes", "no"),
         college = factor(college)) |> 
  filter(income > 0,
         handset_price < 1500) |> 
  select(-id) |> 
  na.omit()

#What other cleanup is needed?

# Check whether the operation was successful
summary(m_clean)

```

### Fit a classification tree and create a confusion matrix 

Fit a tree model and summarize the results in a confusion matrix using the default class decision threshold produced by using the `type = `class"` argument to `predict()`.  Report the counts in the cells.

```{r}
(classification_tree <- rpart(leave ~., 
                              data = m_clean,
                              maxdepth = 5,
                              minbucket = 10))

table(predicted = predict(classification_tree, type = "class"),
      observed = m_clean$leave)

rpart.plot(x = classification_tree)

```

In the MegaTelCo case our objective is to model leave. Correctly predicting leave is therefore a true positive, represented in the upper left cell in the matrix.  Incorrectly predicting leave---we predict leave but the customer stays---is a false positive, in the upper right cell. Typically, in a case like this one we would only consider predictions of leave to be actionable. That is because if a customer is predicted to stay then, logically, no intervention is necessary. We will therefore ignore the second row in this confusion matrix.

### Calculate profit I

Using the confusion matrix in the previous question how much profit (revenue - costs) could be expected with these costs-benefits?

For MegaTelCo we will assume benefit = 800 and cost = 200.  TPs are a benefit, FPs are a cost. Again, we ignore those predicted to stay. (The cost-benefit numbers will be different in the AdviseInvest case! )

Why are TPs a benefit? In the MegaTelCo scenario, these are customers who are predicted to leave and actually were going to leave. If your marketing campaign is successful, then you can convince them to stay, thereby saving the company money. (In the AdviseInvest scenario, TPs are customers that you have predicted will answer the phone and do answer, thus providing an opportunity for your sales reps to make a sale.)


```{r}
table(predicted = predict(classification_tree, type = "class"),
      observed = m_clean$leave)

```

```{r}
#Note: these numbers may have changed
6999 * (800 - 200) - 4579 * 200

```

### Calculate profit II

How much profit (revenue - costs) could be expected if all customers got the marketing incentive? This is equivalent to always predicting leave, so:

```{r}

m_clean |> 
  count(leave)

```

TPs will be the leavers, FPs the stayers.

```{r}
#Note: these numbers may have changed
7470 * (800 - 200) - 7521 * 200

```
 
### Calculate profit III

What if we shifted the class decision threshold down to .2?  What happens to profit?

Here are the steps involved in doing that analysis:

1. predict probabilities

```{r}

predict(classification_tree, type = "prob") |> 
  head()

```

Pick out just the first column using square bracket notation: [row, column]. This is an alternative to dollar sign notation.  [ , column] is the same as $column.

```{r}
predict(classification_tree, type = "prob")[ ,1] |>  # picks out the first column
  head()
```

2. Convert probabilities to a class label at a given decision threshold.

```{r}
ifelse(predict(classification_tree, type = "prob")[ ,1] >= .2, "LEAVE", "STAY")

```

3. Create a confusion matrix using that decision threshold. 

```{r}

table(predicted = ifelse(predict(classification_tree, type = "prob")[ ,1] >= .2, "LEAVE", "STAY"),
      observed = m_clean$leave)


```

4. Calculate profit.  

```{r}
#Note: these numbers may have changed
7470 * (800 - 200) - 5394 *200

```

### Predict for new data and create a customer contact list

Predict the probability of answering for new customers, then filter the list at the optimal class decision threshold, .2. (Why .2?  In an actual analysis this would be one of the calculations you would need to do. Here this optimal decision threshold is provided for you.)

Key idea: The distinction between modeling and using the model. 

So far in the course we have been modeling. The presence of a target variable allows us to create a model that relates predictor variables to the target. In the case of a classification tree, the result is a set of rules that lead to a prediction for customers for whom the target variable has already been observed. But utlimately the reason for modeling is to create a model that can be used --- with new data!  Once we have a model then we are able to  predict either a class label or a probability for customers for whom the target variable has not been observed. For this we use the predict() function. It works just fine without a target variable in the input data set. 

We will create a fake new dataset based on the original, just for illustration, with no target.  (Note: you would never do this in reality.  This code is just for illustration!!!)

```{r}
# For illustration, create new dataset, take out target, and add fake id.  As noted above,
# this is not part of the module 5 project!! For the project this data is already supplied for you
# in adviseinvest_new_customer.csv.

set.seed(123) # The seed guarantees we get exactly the same dataset after random;y subsetting
new_customers <- m_clean |> 
  sample_n(1000) |> 
  select(-leave) |>  
  mutate(id = sample(1:1000, 1000, replace = F))

glimpse(new_customers)

```

Get predictions for current customers:

```{r}
# generate predictions
predict(object = classification_tree, 
        newdata = new_customers, 
        type = "prob")[ ,1] |> 
  head()
```

Add prediction to the dataset and filter for probability >= .2. Select just id and prob for a contact list.

```{r}

contact_list <- new_customers |> 
  mutate(prob_churn = predict(classification_tree, 
                              newdata = new_customers, 
                              type = "prob")[,1]) |> 
  filter(prob_churn >= .2) |> 
  select(id, prob_churn) |> 
  arrange(desc(prob_churn))

contact_list

glimpse(contact_list)

```

##Here there are just 864 out of 1000 customers who should be contacted in order to optimize profit. 

### Recommendation

Make a recommendation based on the analytics you have done. This should be a carefully thought out, well-reasoned, and well-written recommendation in the neighborhood of 2 - 4 paragraphs.  Plus for using tables and plots!

It is critical to think through exactly how the company will use the analytics you have produced. It is your responsibility as the analytics guru to tell them what to do.

